---
title: "Codebook for Konthai"
output: 
  html_document:
    toc: true # table of content true
    toc_float: true
    depth: 3  # upto three depths of headings (specified by #, ## and ###)
    number_sections: true  ## if you want number sections at each table header
    theme: spacelab  # many options for theme, this one is my favorite.
    highlight: tango  # specifies the syntax highlighting style
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

# Executive Summary

This is the codebook for the `konthai` project. It details how these datasets are created:

* `prop_df` The dataset for proportional tests of population characteristics between Thai poeple in the US and the US poopulation

* `county_reg` The dataset for regression analysis with the dependent variable being the number, density and density per area of Thai population and the independent variables being the livability index.

# Load Libraries

```{r library}
library(data.table)
library(sqldf)
```

# Load Datasets

```{r the setup, results='hide'}
input <- 'files/'

#proportion test function
test_prop <- function (x1,x2,n1,n2){
  p_hat <- (x1+x2)/(n1+n2)
  p1 <- x1/n1
  p2<- x2/n2
  numerator <- p1-p2
  denominator <- sqrt(p_hat*(1-p_hat))*sqrt(1/n1 + 1/n2)
  numerator/denominator
}

#livability index AARP Public Policy Institute
livability <- read.csv(paste0(input,'livability.csv'))

#LANGUAGE SPOKEN AT HOME BY ABILITY TO SPEAK ENGLISH FOR THE POPULATION 5 YEARS AND OVER Universe: Population 5 years and over  more information 2010-2014 American Community Survey 5-Year Estimates
ACS_14_5YR_B16001 <- read.csv(paste0(input,'ACS_14_5YR_B16001/ACS_14_5YR_B16001_with_ann.csv')
                              , skip = 1)

#ASIAN ALONE BY SELECTED GROUPS Universe: Total Asian alone population  more information 2010-2014 American Community Survey 5-Year Estimates
ACS_14_5YR_B02006 <- read.csv(paste0(input,'ACS_14_5YR_B02006/ACS_14_5YR_B02006_with_ann.csv')
                              , skip = 1)

#SELECTED POPULATION PROFILE IN THE UNITED STATES  more information 2014 American Community Survey 1-Year Estimates
ACS_14_1YR_S0201 <- read.csv(paste0(input,'ACS_14_1YR_S0201/ACS_14_1YR_S0201_with_ann.csv')
                             , skip = 1)

#SELECTED POPULATION PROFILE IN THE UNITED STATES  more information 2014 American Community Survey 1-Year Estimates for Thai people
ACS_14_1YR_S0201_thai <- read.csv(paste0(input,'ACS_14_1YR_S0201_thai/ACS_14_1YR_S0201_with_ann.csv')
                             , skip = 1)

```

# For the Proportional Tests

Our goal is to perform a series of proportional z-tests between population characteristics of Thai people in the US and the total US population. In order to do that, we transform data from `A2014 American Community Survey 1-Year Estimates` into our desired form `prop_df`. This resulted in 141 variables in 45 categories.

## Transformation

```{r the transformation thai_overall}
thai_overall <- transpose(ACS_14_1YR_S0201_thai)

#remove ids
thai_overall <- data.frame(thai_overall[6:nrow(thai_overall),])


#get sensible variable names
thai_overall$var <- sapply(colnames(ACS_14_1YR_S0201_thai[,6:ncol(ACS_14_1YR_S0201_thai)]), 
                           FUN= function(x) tail(strsplit(x, split='...',fixed=TRUE)[[1]],1))

#get category
thai_overall$cat <- sapply(colnames(ACS_14_1YR_S0201_thai[,6:ncol(ACS_14_1YR_S0201_thai)]), 
                           FUN= function(x) tail(strsplit(x, split='...',fixed=TRUE)[[1]],2)[1])

#get detailed category
thai_overall$d_cat <- sapply(colnames(ACS_14_1YR_S0201_thai[,6:ncol(ACS_14_1YR_S0201_thai)]), 
                           FUN= function(x) 
                              paste(
                              head(strsplit(x, split='...',fixed=TRUE)[[1]],-1)
                              , collapse = '_')
                          )

#remove error margins
thai_overall$iserror <- sapply(colnames(ACS_14_1YR_S0201_thai[,6:ncol(ACS_14_1YR_S0201_thai)]), 
                           FUN= function(x) grepl(pattern = 'Margin.of.Error',x))
thai_overall <- thai_overall[thai_overall$iserror==FALSE,]

#make columns look nice
colnames(thai_overall) <- c('val','var','cat','d_cat','iserror')
thai_overall <- thai_overall[c('d_cat','cat','var','val')]
#parse to numeric
thai_overall$val <- as.numeric(as.character(thai_overall$val))

#separate percentage and total
thai_overall$istotal <- sapply(thai_overall$cat, 
                           FUN= function(x) grepl(pattern = 'Estimate',x))

#percentage
thai_per <- thai_overall[!thai_overall$istotal,]

#total
thai_total <- thai_overall[thai_overall$istotal & thai_overall$val > 1000,]

#join percentage and total
thai_pertal <- sqldf("
select
  thai_per.d_cat,
	thai_per.cat,
	thai_per.var,
	thai_per.val as percentage,
	thai_total.val as total
from thai_per
inner join thai_total
on thai_per.cat = thai_total.var
                   ")
#remove duplicates
thai_pertal <- thai_pertal[!duplicated(thai_pertal),]

#calulate number based on percentage
thai_pertal$num <- (thai_pertal$percentage/100) * thai_pertal$total
```


```{r the transformation us_overall}
us_overall <- transpose(ACS_14_1YR_S0201)

#remove ids
us_overall <- data.frame(us_overall[6:nrow(us_overall),])


#get sensible variable names
us_overall$var <- sapply(colnames(ACS_14_1YR_S0201[,6:ncol(ACS_14_1YR_S0201)]), 
                           FUN= function(x) tail(strsplit(x, split='...',fixed=TRUE)[[1]],1))

#get category
us_overall$cat <- sapply(colnames(ACS_14_1YR_S0201[,6:ncol(ACS_14_1YR_S0201)]), 
                           FUN= function(x) tail(strsplit(x, split='...',fixed=TRUE)[[1]],2)[1])

#get detailed category
us_overall$d_cat<-sapply(colnames(ACS_14_1YR_S0201[,6:ncol(ACS_14_1YR_S0201)]), 
                           FUN= function(x) 
                              paste(
                              head(strsplit(x, split='...',fixed=TRUE)[[1]],-1)
                              , collapse = '_')
                          )

#remove error margins
us_overall$iserror <- sapply(colnames(ACS_14_1YR_S0201[,6:ncol(ACS_14_1YR_S0201)]), 
                           FUN= function(x) grepl(pattern = 'Margin.of.Error',x))
us_overall <- us_overall[us_overall$iserror==FALSE,]

#make columns look nice
colnames(us_overall) <- c('val','var','cat','d_cat','iserror')
us_overall <- us_overall[c('d_cat','cat','var','val')]
#parse to numeric
us_overall$val <- as.numeric(as.character(us_overall$val))

#separate percentage and total
us_overall$istotal <- sapply(us_overall$cat, 
                           FUN= function(x) grepl(pattern = 'Estimate',x))

#percentage
us_per <- us_overall[!us_overall$istotal,]

#total
us_total <- us_overall[us_overall$istotal & us_overall$val > 1000,]

#join percentage and total
us_pertal <- sqldf("
select
  us_per.d_cat,
	us_per.cat,
	us_per.var,
	us_per.val as percentage,
	us_total.val as total
from us_per
inner join us_total
on us_per.cat = us_total.var
                   ")
#remove duplicates
us_pertal <- us_pertal[!duplicated(us_pertal),]

#calulate number based on percentage
us_pertal$num <- (us_pertal$percentage/100) * us_pertal$total
```


```{r join thai_pertal and us_pertal}

prop_df <- sqldf("
select
  us_pertal.d_cat,
	us_pertal.cat,
	us_pertal.var,
  us_pertal.percentage as us_percentage,
	us_pertal.num as us_num,
	us_pertal.total as us_total,
  thai_pertal.percentage as thai_percentage,
	thai_pertal.num as thai_num,
	thai_pertal.total as thai_total
from us_pertal
inner join thai_pertal
on us_pertal.cat = thai_pertal.cat and
	us_pertal.var = thai_pertal.var
                 ")
prop_df$z_score <- test_prop(prop_df$thai_num, prop_df$us_num, 
                             prop_df$thai_total, prop_df$us_total)

prop_df <- prop_df[!is.nan(prop_df$z_score),]

prop_df <- prop_df[c('d_cat','var','thai_percentage','thai_num','thai_total',
                     'us_percentage','us_num','us_total','z_score')]
#save to csv
write.table(prop_df, file = paste(input,"prop_df.csv",sep=""), 
              sep="\t",row.names=FALSE,fileEncoding = "UTF-8",quote = TRUE)

```

## Variable Explanation

* `d_cat` Detailed category of the variable

```{r categories}
unique(prop_df$d_cat)
```

* `var` Variable such as male and female
* `thai_percentage` Percentage of population in the variable for Thai population in the US
* `thai_num` Number of Thai population in the US for the variable
* `thai_total` Total number of Thai population in the US for the category
* `us_percentage` Percentage of population in the variable for US population
* `thai_num` Number of US population for the variable
* `thai_total` Total number of US population for the category
*  `z-score` Z-score for proportional test

# For the Regression Analysis

We combine the `livability` dataset obtained by the script `extract.ipynb` with `Total Asian alone population` and `LANGUAGE SPOKEN AT HOME BY ABILITY TO SPEAK ENGLISH` from `2010-2014 American Community Survey 5-Year Estimates` to get the `county_reg` dataset.

```{r}

county_reg <-data.frame(
  county = ACS_14_5YR_B02006$Geography,
  total = ACS_14_5YR_B02006$Estimate..Total.,
  thai_num=ACS_14_5YR_B02006$Estimate..Total....Thai,
  thai_per= ACS_14_5YR_B02006$Estimate..Total....Thai / ACS_14_5YR_B02006$Estimate..Total.,
  thai_vwell=ACS_14_5YR_B16001$Estimate..Total....Thai....Speak.English..very.well.,
  thai_vwell_per=ACS_14_5YR_B16001$Estimate..Total....Thai....Speak.English..very.well. /
    ACS_14_5YR_B02006$Estimate..Total....Thai,
  thai_less=ACS_14_5YR_B16001$Estimate..Total....Thai....Speak.English.less.than..very.well.,
    thai_less_per=ACS_14_5YR_B16001$Estimate..Total....Thai....Speak.English.less.than..very.well.
    / ACS_14_5YR_B02006$Estimate..Total....Thai
)

county_reg <- cbind(county_reg,
                    subset(livability, select = -c(county)))

#save to csv
write.table(county_reg, file = paste(input,"county_reg.csv",sep=""), 
              sep="\t",row.names=FALSE,fileEncoding = "UTF-8",quote = TRUE)

```

## Variable Explanation

* `count` County name
* `total` Total population of the county
* `thai_num` Number of Thai people
* `thai_num_per` Percentage of Thai people over total population
* `thai_vwell` Number of Thai people who speak English very well
* `thai_vwell_per` Percentage of Thai people who speak English very well
* `thai_less` Number of Thai people who speak English less than very well
* `thai_less_per` Percentage of Thai people who speak English less than very well
* `engagement`, `environment`, `health`, `housing`, `neighborhood`, `opportunity`, and `transportation` Scoring variables out of 100 from [AARP Livability Index](http://www.aarp.org/ppi/issues/livable-communities/info-2015/livability-index.html)
